Turbo fast no compromise gravitational wave parameter estimation

### 2022-11-02 src/tex/ms.tex
upload
Adding normalizing flow paragraphs

@@

The computational cost of evaluating a waveform model scales linearly with the number of sample points either in time or frequency domain.
This introduces greater computational burden for longer-duration signals.
To reduce the computational cost, there are a number of methods to reduce the number of basis points one would need to compute the likelihood faithfully \kw{cite}.
We use heterodyned likelihood \kw{cite} (also named relative binning in \kw{cite}) in this work.

@@

While

@@

% Higher dimensional problems
While standard GW analysis goes up to 17 dimension, 
\begin{enumerate}
    \item Multi-device scaling?
    \item Other detectors configuration
    \item  Combining with noise inference
    \item Higher dimensional problems need HMC still
\end{enumerate}

>>>>

\begin{align}
    p(\theta| d) = \frac{\mathcal{L}(d|\theta)p(\theta)}{p(d)}
\end{align}

\begin{align}
    \log{\mathcal{L}(d|\theta)} = -\left<d-h|d-h\right>/2
\end{align}

\begin{align}
    \left<a|b\right> = 4 Re\Int \frac{a(f)^*b(f)}{\mathcal{S}_n(f)} df
\end{align}

@@

The computational cost of evaluating a waveform model scales linearly with the
number of sample points either in time or frequency domain. This introduces
greater computational burden for longer-duration signals. To reduce the
computational cost, there are a number of methods to reduce the number of basis
points one would need to compute the likelihood faithfully \kw{cite}. We use
heterodyned likelihood \kw{cite} (also named relative binning in \kw{cite}) in
this work. We give a concise description of what we implement in our code here.
For more extensive discussion of the derivation of heterodyned likelihood, we
refer the reader the reference \kw{}.

@@

While gradient-based samplers have been shown to be superior in terms of
performance when compared to other gradient-free algorithms in many practical
examples\kw{cite}, there are still a number of classes of problems
gradient-based samplers do not solve well. \footnote{To be specific, we are
referring to sampling algorithms that use first order derivative here. Sampling
algorithms that use information of higher order derivative such as manifold-MALA
and Riemannian-HMC can in principle decorrelate local correlation in the target
distribution, however they often have instability issue when they are used on
real-life application, so they are not used often in practice.} For examples,
target distribution that exhibits local correlation are hard to dealt with,
since by construction first order gradient-based algorithms can only handle
global correlation structure. Another example is multi-modality. If there are
multiple modes in the target distribution, an independent chain will likely to
be trapped in one mode and take extremely long time to transverse between the
modes. This means the relative weights between modes will much longer to sample
compared to the shape of each mode.

Normalizing flow is a neural-network based technique that aims at learning a
mapping between a simple distribution, such as a Gaussian distribution, to a
complex distribution, often given in the form of data samples from a target
distribution. Once the network is trained, one can evaluate the probability
density of the complex distribution and samples from it very efficiently.

\begin{align}
    p(x) = p(z) \prod_i \left| \frac{\partial f}{\partial x_i}\right|
\end{align}

For a detail discussion of the algorithm, we refer the readers to \kw{cite}.

@@

% Higher dimensional problems
While standard GW analysis goes up to 17 dimensions,

% Compared to other works

There are recent works from different groups on speeding up parameter estimation
of gravitational wave, including efficient reparameterization and deep learning
\kw{cite}. While all of these methods are pushing down minutes-scale parameter
estimation with high fidelity, we would like to highlight the unique strength of
this work. Compared to \kw{Stephan's paper}, we do not rely on pre-training the
neural network on a large collection of waveforms and noise realization. This
means as soon as new waveforms models and noise models are available, our
algorithm can already be deployed. Furthermore, our method is essentially an
MCMC algorithm, which inherits the merit of convergence measures in MCMC. As we
are only using the normalizing flow as a proposal distribution, and the
normalizing flow is trained jointly with a local sampler, we do not suffer from
overfitting since our data is always approaching the target distribution, in
this sense, we do not introduce potential extra systematic error to inference
result. Compared to \kw{Tousif's paper}, we do not rely on handcrafted
reparameterization of the coordinate systems. If the reparameterization scheme
is known ahead of time, a nice reparameterization is also encouraged. However,
handcrafted reparameterization depends on the assumptions used in deriving the
reparameterization scheme, which would inevitably run into limitations of use
cases. Our work can be viewed as an automatic reparameterization powered by
normalizing flow, while not being exact hence as efficient as an explicit
reparameterization, the method proposed in this work is applicable to a much
boarder class of problems, such as parameter estimation with precessing
waveforms, testing GR, and multi-event joint inference.

\begin{enumerate}
    \item Multi-device scaling?
    \item Other detectors configuration
    \item  Combining with noise inference
    \item Higher dimensional problems need HMC still
\end{enumerate}
